### 2016-05-03 writing/latex/introduction.tex
Make Edits. Temp remove code because borked computer
Add definition of mpcontainment by Anders, add subpattern def

We can display a permutation on a \emph{figure} in order to give a graphical
representation of the permutation. In such a figure we plot the points
\((i,\pi(i))\) in a Cartesian coordinate system. The figure of the permutation
\(\pi = \perm{2,3,1}\) is shown below
\begin{equation*}
    \pattern{}{2,3,1}{}
\end{equation*}
It is convenient to call the elements of the permutation \emph{points} when
referring to these figures.

@@

\section{Mesh Patterns}
Mesh patterns were introduced by \textcite{journals/combinatorics/BrandenC11} to
capture explicit expansions for certain permutation statistics. They are a
natural extension of classical permutation patterns. A \emph{mesh pattern} is a pair
\begin{equation*}
    p = (\tau,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
    [0,k]\times [0,k].
\end{equation*}

The set \(R\) is called the \emph{mesh set} of the mesh pattern \(p\).
The figure for a mesh pattern looks similar to that of a classical pattern with the
addition that we shade the unit square with bottom left corner \((i,j)\) for each \((i,j) \in R\):
\begin{equation*}
    \pattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }
\end{equation*}

We define containment, and avoidance, of the pattern \(p\) in the permutation
\(\pi\) on mesh patterns analogously to classical containment, and avoidance,
of \(\tau\) in \(\pi\) with the additional restrictions on the relative
position of the occurence of \(\tau\) in \(\pi\). These restrictions say that
no elements of \(\pi\) are allowed in the regions of the figure corresponding
to shaded boxes in the mesh (see \cite[2]{journals/combinatorics/BrandenC11}
for the precise definition).

@@

Let us consider the figure for the permutation \(\pi\). The subsequence \(325\) is
an occurrence of \(213\) in the classical sense and the remaining points of \(\pi\)
are not contained in the regions corresponding to the shaded boxes in \(p\).
\begin{equation*}
    \pattern{1,3,5}{3,4,2,1,5}{0/2,0/3,0/4,
                               1/0,1/1,1/2,
                               2/0,2/1,2/2,
                               3/2,3/3,3/4,
                               4/2,4/3,4/4}
\end{equation*}
The subsequence \(325\) is therefore an occurence of the pattern \(p\) in \(\pi\) and
\(\pi\) contains \(p\).

Now we consider the figure for the permutation \(\sigma\). This permutation
avoids \(p\) since for every occurence of the classical pattern \(\perm{2,1,3}\)
there is at least one point in one of the shaded boxes. For example, consider the
subsequence\(315\) in \(\sigma\), this is an occurence of \(\perm{2,1,3}\) but
not the mesh pattern since the points with values \(4\) and \(2\) are in the
regions corresponding to the boxes \((0,1)\) and \((0,2)\), which are shaded in
\(p\). This is shown in the figure below.

@@

>>>>

We can display a permutation in a \emph{plot} in order to give a graphical
representation of the permutation. In such a plot we display the points
\((i,\pi(i))\) in a Cartesian coordinate system. The plot of the permutation
\(\pi = \perm{2,3,1}\) is shown below
\begin{equation*}
    \pattern{}{2,3,1}{}
\end{equation*}
It is convenient to call the elements of the permutation \emph{points} when
referring to these plots.

@@

\section{Mesh Patterns}
Mesh patterns were introduced by \textcite{journals/combinatorics/BrandenC11} to
capture explicit expansions for certain permutation statistics. They are a
natural extension of classical permutation patterns.
A \emph{mesh pattern} is a pair
\begin{equation*}
    p = (\pi,R)\text{ with } \tau \in \mathfrak{S}_k \text{ and } R \subseteq
    [0,k]\times [0,k].
\end{equation*}

The set \(R\) is called the \emph{mesh set} of the mesh pattern \(p\).
The plot for a mesh pattern looks similar to that of a classical pattern with the
addition that we shade the unit square with bottom left corner \((i,j)\) for each \((i,j) \in R\):
\begin{equation*}
    \pattern{}{ 2, 1, 3 }{ 0/1, 0/2, 1/0, 1/1, 2/1, 2/2 }
\end{equation*}

We define containment, and avoidance, of the pattern \(p\) in the permutation
\(\tau\) on mesh patterns analogously to classical containment, and avoidance,
of \(\pi\) in \(\tau\) with the additional restrictions on the relative
position of the occurence of \(\pi\) in \(\tau\). These restrictions say that
no elements of \(\tau\) are allowed in the regions of the plot corresponding
to shaded boxes in the mesh. Formally defined by \textcite{journals/combinatorics/BrandenC11},
an \emph{occurrence} of \(p\) in \(\tau\) is a subset \(\omega\) of the plot of
\(\tau, G(\tau) = \{(i,\tau(i)\mid i\in[1,n]\}\) such that there are order-preserving
injections \(\alpha,\beta:[1,k]\mapsto[1,n]\) satisfying the following two conditions.

\noindent Firstly, \(\omega\) is an occurrence of \(\pi\) in the classical sense
  \begin{enumerate}[i.]
    \item \(\omega = \{(\alpha(i),\beta(j)):(i,j)\in G(\pi)\}\)
  \end{enumerate}\saveenum
  Define \(R_{ij} = [\alpha(i)+1,\alpha(i+1)-1]\times[\beta(j)+1,\beta(j+1)-1]\)
   for \(i,j\in[0,k]\) where \(\alpha(0)=\beta(0)=0\) and \(\alpha(k+1)=\beta(k+1)=n+1\). Then
   the second condition is
  \begin{enumerate}[i.]\resetenum
  \item if \((i,j) \in R \text{ then } R_{ij} \cap G(\tau) = \emptyset\)
\end{enumerate}
We call \(R_{ij}\) the region corresponding to \((i,j)\).
We define containment of a mesh pattern \(p\) in another mesh \(\kappa\) as above,
with the additional condition that if \((i,j)\in R\text{ then } R_{ij}\) is
contained in the mesh set of \(\kappa\), in this case we call \(p\) a \emph{subpattern}
of \(\kappa\).

@@

Let us consider the plot for the permutation \(\pi\). The subsequence \(325\) is
an occurrence of \(213\) in the classical sense and the remaining points of \(\pi\)
are not contained in the regions corresponding to the shaded boxes in \(p\).
\begin{equation*}
    \pattern{1,3,5}{3,4,2,1,5}{0/2,0/3,0/4,
                               1/0,1/1,1/2,
                               2/0,2/1,2/2,
                               3/2,3/3,3/4,
                               4/2,4/3,4/4}
\end{equation*}
The subsequence \(325\) is therefore an occurence of the pattern \(p\) in \(\pi\) and
\(\pi\) contains \(p\).

Now we consider the plot for the permutation \(\sigma\). This permutation
avoids \(p\) since for every occurence of the classical pattern \(\perm{2,1,3}\)
there is at least one point in one of the shaded boxes. For example, consider the
subsequence\(315\) in \(\sigma\), this is an occurence of \(\perm{2,1,3}\) but
not the mesh pattern since the points with values \(4\) and \(2\) are in the
regions corresponding to the boxes \((0,1)\) and \((0,2)\), which are shaded in
\(p\). This is shown in the figure below.

@@

Two patterns are said to be \emph{coincident} if they avoid the same set of
permutations and \emph{Wilf-equivalent} if they avoid the same number of permutations
at every length.

### 2016-05-03 writing/latex/wilf.tex
Add definition of mpcontainment by Anders, add subpattern def

\nextvar
The pattern \(p_1 = \textpattern{}{1}{0/1,1/0}\) is one of the
quadrant marked mesh patterns studied by Remmel,
Kitaev and Tiefenbruck\cite{2012arXiv1201.6243K}. Alternatively we can enumerate
avoiders by decomposing a typical avoider of \(p_1\) around the maximum element
in order to give the following structural decomposition.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup \decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
Since if the upper right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p_1\). The lower left region however can create occurrences
of \(p_1\) and therefore must also avoid \(p_1\) as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying.
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1+x-xC(x)}
\end{align*}

@@

\nextvar[\varfine]
It can be shown by use of \PropositionRef{prop:dom2} that the patterns
\textpattern{}{1}{0/0,1/1} and \(p_2 = \textpattern{}{1}{0/0,0/1,1/1}\) are
coincident. Consider the decomposition of an avoider of \(p_2\) in
\(\av{\perm{2,3,1}}\) around the maximum
\begin{equation*}
    \scriptvar_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
This can be explained succinctly by the fact that a permutation containing
\(p_2\) starts with it's maximum, by not allowing the left part of the
\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever happening.

Consider \(p_3 = \textpattern{}{1}{0/1,1/0,1/1}\), avoiding this pattern
means that a permutation does not end with it's maximum. We can perform a similar
decomposition as before to get
\begin{equation*}
    \mathcal{\var}_2 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}

Now consider \(p_4 = \textpattern{}{1}{0/0,0/1,1/0}\), the avoiders of this
pattern are permutations that do not start with their minimum. In this case we
perform the decomposition around the leftmost element
\begin{equation*}
    \mathcal{\var}_3 =\varepsilon \sqcup \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
All of these classes have the same generating function namely
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1)
\end{equation*}
Enumeration of this generating function gives
\begin{equation*}
    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc
\end{equation*}

\nextvar[\varmaxl]
There is one pattern of length \(1\) still to consider. The pattern
\(p_5 = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
that do not end in their minimum. Considering the standard decomposition
of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
of \(p_5\) must fit into the following form precisely once.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup
    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{
        Minimum comes before the maximum.
    } \sqcup
    \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\scriptstyle{\scriptvar\setminus\varepsilon}}}_\text{
        Minimum is after the maximum, cannot be last.
    }
\end{equation*}
Therefore this particular class has generating function \(\var(x)\) satisfying
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1) + x(\var(x)-1)
\end{equation*}
This generating function can be enumerated to give
\begin{equation*}
    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc
\end{equation*}

>>>>

\nextvar
The pattern \(p = \textpattern{}{1}{0/1,1/0}\) is one of the
quadrant marked mesh patterns studied by Remmel,
Kitaev and Tiefenbruck\cite{2012arXiv1201.6243K}. Alternatively we can enumerate
avoiders by decomposing a typical avoider of \(p_1\) around the maximum element
in order to give the following structural decomposition.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup \decompmax{0/1,1/0}{\scriptvar}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}
Since if the upper right section was empty the maximum would create an
occurrence of the pattern, however no points in this section can create
an occurrence since the maximum lies in a region corresponding to the
shading in \(p\). The lower left region however can create occurrences
of \(p\) and therefore must also avoid \(p\) as well as \(\perm{2,3,1}\).
This gives the generating function of avoiders to be the function \(\var(x)\)
satisfying.
\begin{align*}
    \var(x) &= 1 + x\var(x)(C(x)-1) \\
    \shortintertext{Solving for \(\var\) gives}
    \var(x) &= \frac{1}{1+x-xC(x)}
\end{align*}

@@

\nextvar[\varfine]
It can be shown by use of \PropositionRef{prop:dom2} that the patterns
\textpattern{}{1}{0/0,1/1} and \(q_1 = \textpattern{}{1}{0/0,0/1,1/1}\) are
coincident. Consider the decomposition of an avoider of \(q_1\) in
\(\av{\perm{2,3,1}}\) around the maximum
\begin{equation*}
    \scriptvar_1 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
This can be explained succinctly by the fact that a permutation containing
\(q_1\) starts with it's maximum, by not allowing the left part of the
\(\perm{2,3,1}\) avoider to be empty we prevent an occurrence from ever happening.

Consider \(q_2 = \textpattern{}{1}{0/1,1/0,1/1}\), avoiding this pattern
means that a permutation does not end with it's maximum. We can perform a similar
decomposition as before to get
\begin{equation*}
    \mathcal{\var}_2 =\varepsilon \sqcup \decompmax{0/1,1/0}{\mathcal{C}}{}{}{\mathcal{C}\setminus\varepsilon}
\end{equation*}

Now consider \(q_3 = \textpattern{}{1}{0/0,0/1,1/0}\), the avoiders of this
pattern are permutations that do not start with their minimum. In this case we
perform the decomposition around the leftmost element
\begin{equation*}
    \mathcal{\var}_3 =\varepsilon \sqcup \decompleft{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}
\end{equation*}
All of these classes have the same generating function namely
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1)
\end{equation*}
Enumeration of this generating function gives
\begin{equation*}
    1, 0, 1, 3, 9, 28, 90, 297, 1001, 3432, 11934,\dotsc
\end{equation*}

\nextvar[\varmaxl]
There is one pattern of length \(1\) still to consider. The pattern
\(r = \textpattern{}{1}{0/0,1/0,1/1}\) is avoided by all permutations
that do not end in their minimum. Considering the standard decomposition
of a \(\perm{2,3,1}\) avoider around the maximum we can see that an avoider
of \(r\) must fit into the following form precisely once.
\begin{equation*}
    \scriptvar = \varepsilon \sqcup
    \underbrace{\decompmax{0/1,1/0}{\mathcal{C}\setminus\varepsilon}{}{}{\mathcal{C}}}_\text{
        Minimum comes before the maximum.
    } \sqcup
    \underbrace{\decompmax{0/0,0/1,1/0}{}{}{}{\scriptstyle{\scriptvar\setminus\varepsilon}}}_\text{
        Minimum is after the maximum, cannot be last.
    }
\end{equation*}
Therefore this particular class has generating function \(\var(x)\) satisfying
\begin{equation*}
    \var(x) = 1 + xC(x)(C(x)-1) + x(\var(x)-1)
\end{equation*}
This generating function can be enumerated to give
\begin{equation*}
    	1, 0, 1, 4, 13, 41, 131, 428, 1429, 4861, 16795,\dotsc
\end{equation*}
